Optimizing Automated Finite Element Integration through Expression Rewriting and Code Specialization

### 2014-08-06 TOMS_Paper/assembly.tex
Rewrite lots of stuff...

\begin{figure}[t]
\tiny
\centering     
\subfigure[Original code]{\label{fig:original-code}\lstinputlisting{listings/simple.code}}
~~~~~~~~~~~~~
\subfigure[Factorized code]{\label{fig:factorized-code}\lstinputlisting{listings/factorized.code}}
\caption{Original and factorized code.}
\end{figure}

Consider the simplified example of the element matrix computation in Figure~\ref{fig:original-code}, which is an excerpt from a real Burgers problem. In practice, depending on the problem, the expression tree could be much more complex, with multiple levels of nesting. The example is however representative for a large class of problems, so we will use it throughout the rest of the paper for illustrative purpose.

A first glimpse of the code suggests that the sub-expression \texttt{F[i][j]*d + G[i][j]*e} is invariant with respect to the innermost loop, so it should be hoisted at the level of the outer loop \texttt{j}. This is a standard compiler transformation, which is supported by any available compilers. With a closer look we notice that the sub-expression \texttt{a*C[i][k] + b*E[i][k]} is also invariant, although, in this case, with respect to the outer loop \texttt{j}. In~\cite{Luporini}, we have showed that a \textit{generalized} loop-invariant code motion transformation - that is, given a non-trivial expression, the capability of analyzing all of its sub-expressions with respect to the enclosing loops to determine what code is hoistable- is not supported by available compilers. Moreover, the lack of cost models to ascertain both the optimal place where to hoist an expression and whether or not vectorizing it at the price of extra temporary memory is a fundamental limiting factor. We have addressed these problems by implementing a generalized loop-invariant code motion transformation in COFFEE.

We now consider the case of forms that ``hide'' further opportunities for code hoisting. Such forms are by no means exotic: for example, the pattern described next is commonly found in elastic problems. By examining again the example in Figure~\ref{fig:original-code}, we notice that the basis function array \texttt{B}, iterating along the iteration space \texttt{[i,j]}, appears twice in the expression. If we expand the products in which \texttt{B} is accessed, we can apply product commutativity and then factorize the expression as in Figure~\ref{fig:factorized-code}. This has two effects: firstly, it reduces the total number of arithmetic operations performed; secondly, and most importantly, it exposes a new sub-expression \texttt{(A[i][k]/c + T2[k]*f)} invariant with respect to loop \texttt{j}. Therefore, code hoisting can be performed.

The second observation we make concerns the register pressure induced by the expression. Once loop-invariant terms are lifted, we can think about data locality and, in particular, register allocation. Assume the local assembly kernel is executed on a state-of-the-art architecture having 16 logical registers, e.g. an Intel Haswell. Each value appearing in the expression is loaded and kept in a register as long as possible. For instance, the scalar value \texttt{g} is loaded once, whereas the term \texttt{det*W[i]} is precomputed and loaded in a register at every $i$ iteration. This implies that at every iteration of the \texttt{jk} loop nest, 12$\%$ of the available registers are spent just to store constant values. In more complicated expressions, the percentage of registers destined to store loop-invariant terms can be even higher. Registers are, however, a precious resource, especially when evaluating intensive expressions. The smaller is the number of free registers, the worse is the instruction-level parallelism achieved: for example, a shortage of registers can increase the pressure on the L1 cache (i.e. it can worsen data locality), or it may prevent the effective application of standard transformations like loop unrolling. The ER works around this problem by suitably expanding terms and introducing, where necessary, new temporary values.

@@

\begin{figure}
\tiny
\lstinputlisting{listings/toexpand.code}
\caption{Expandable code}\label{fig:toexpand-code}
\end{figure}

Consider the variant of the running (transformed) example shown in Figure~\ref{fig:toexpand-code}. Again, this is a representative example of what happens in real finite element forms. We can easily distribute \texttt{det*W[i]} over the three operands on the left-hand side of the multiplication, and then absorb it in the pre-computation of \texttt{T1}, resulting in the code illustrated in Figure~\ref{fig:expanded-1-code}. Freeing the register destined to the constant \texttt{g} is more complicated: we cannot absorb it in the pre-computation of \texttt{T1} because the same array is accessed in the evaluation of \texttt{(T1[j]*A[i][k])}. The solution is to add another temporary as in Figure~\ref{fig:expanded-2-code}. Generalizing, this is a problem of data dependencies; in order to solve it, we employ a dependency graph in which we add a direct edge from identifier \texttt{A} to identifier \texttt{B} to denote that the evaluation of \texttt{B} depends on \texttt{A}. The dependency graph is initially empty; every time a new temporary is created due to loop-invariant code motion or expansion of terms is performed, it is updated by suitably adding vertices and edges.

\begin{figure}[t]
\tiny
\centering     
\subfigure[Expanded 1 code]{\label{fig:expanded-1-code}\lstinputlisting{listings/expanded-1.code}}
~~
\subfigure[Expanded 2 code]{\label{fig:expanded-2-code}\lstinputlisting{listings/expanded-2.code}}
\caption{Expanded code.}\label{fig:expanded-code}
\end{figure}

\subsection{Rewrite Rules}
In general, assembly expressions produced by automated code generation can be much more complex (more terms and operations involved) and nested. Our goal is to establish a portable, platform- and compiler-independent, and systematic way of reducing the strength of an expression. The technique should be simple; definitely it must be robust to be integrated in an optimizing domain-specific compiler capable of supporting real problems. Ideally, it should be naturally extendible to problems that will be supported in next releases of state-of-the-art frameworks like Firedrake and FEniCS: for instance, explicit support for outer-product finite elements will enable generation of kernels with much deeper loop nests, and the ER should transparently be able to deal with these structures as well.

To address these issues, we have based the implementation of the ER in COFFEE on a set of formal rewrite rules. By applying these rules, it is possible to derive how an expression will be transformed, as well as what and where (i.e. at which level in the loop nest) temporaries will be introduced. When applying a rule, the ER needs to update the state of the loop nest, to reflect, for example, the use of a new temporary and the newly created data dependencies. We define, therefore, the state of a loop nest $L = (\sigma, G)$, where $G = (V, E)$ represents the dependency graph, while $\sigma$ maps invariant sub-expressions to identifiers of temporary arrays. The notation $\sigma_i$ refers to invariants hoisted at the level of loop $i$. We also introduce the \textit{conditional hoister} operator $[]$ on $\sigma : Inv \rightarrow S$ such that
\begin{gather*}
\sigma[\sfrac{v}{x}] = 
\begin{cases}
\sigma(x) \text{~~~~~~~~if $x \in Inv$; $v$ is ignored}\\
v \text{~~~~~~~~~~~~~if $x$ $\notin$ $Inv$; $\sigma$(x) = v}\\
\end{cases}
\end{gather*}
That is, intuitively, if the invariant expression \texttt{x} has already been hoisted, then return the temporary identifiers that hosts its value; otherwise, hoist the expression. There is a special case when $v = \perp$, used to delete entries in $\sigma$. Specifically:
\begin{gather*}
\sigma[\sfrac{\perp}{x}] = 
\begin{cases}
\sigma(x) \text{~~~~~~~~if $x \in Inv$; $\sigma = \sigma \setminus (x, \sigma(x))$}\\
t \text{~~~~~~~~~~~~~~if $x$ $\notin$ $Inv$; $t \notin Inv$}\\
\end{cases}
\end{gather*}
In other words, the previously hoisted expression \texttt{x} is removed (if any) and the temporary identifier that was hosting its value is returned. This is useful to express updates of invariant expressions.

A generic (sub-)expression is represented with roman letter \texttt{a, b, ...}; constant terms are considered a special case, so greek letters \texttt{$\alpha$, $\beta$} are used instead. The iteration vector $i = [i_0, i_1, ...]$ is the ordered sequence of the indices of the loops enclosing an (sub-)expression. We will refer to $i_0$ as the outermost enclosing loop. The notation $a_i$, therefore, indicates that the expression $a$ assume distinct values while iterating along the loops in $i$; its outermost loop is silently assumed to be $i_0$.

Rewrite rules for the ER are provided in Figure~\ref{fig:rewrite-rules}; obvious rules are omitted for brevity. Conceptually, the ER visits the expression tree from the root, which is the outermost operation, and applies the transformations dictated by the rewrite rules. As an example, one can try instantiating the rules in the code of Figures~\ref{fig:original-code} and~\ref{fig:toexpand-code}; eventually, the optimized code in Figures~\ref{fig:factorized-code} and~\ref{fig:expanded-2-code} is obtained, respectively.

@@

... TODO... : each level in sigma is then taken and wrapped in a loop

...TODO...: to what extent applying rewrite rules, is problem specific...

...TODO...: things get interesting if the kernel operates on a batch of elements...can save a LOT of flops
% blablabla molto complesse, nestings etc expose etc allora rewrite rules

% we will show the effects of this stuff over plain LICM in section

% hoisting constants release registers!!!!

@@

\subsection{Vector-expansion of Invariant Terms}
\label{sec:vect-exp}
The Expression Rewriter's loop-invariant code motion hoists invariant sub-expressions in the body of the outermost loop along which they iterate. If such a sub-expression computes a value depending on the outermost loop only, then the only way of vectorizing it -- if we exclude superword level parallelism~\cite{SLP}, which is in general not applicable to our kernels -- is by vector-expansion of terms.

\begin{figure}[t]
\tiny
\centering     
\subfigure[To vector expand code]{\label{fig:tovectexpand-code}\lstinputlisting{listings/tovectexpand.code}}
~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\subfigure[Vector-expanded code]{\label{fig:vectexpanded-code}\lstinputlisting{listings/vectexpanded.code}}
\caption{Before and after vector expansion}\label{fig:skip-code}
\end{figure}

An example is given in Figure~\ref{fig:tovectexpand-code}: \texttt{F1, F2, F3, T1} are all scalar quantities, and their computation will not be auto-vectorized (as explained in Section~\ref{sec:perf-results}, we experimented with the Intel and GNU compilers). The argument that the cost of evaluating such expressions is not relevant because their computational cost is $O(I)$, while the evaluation of the element matrix has a cost of $O(I T^2)$, is too superficial, in general. There are notable cases, for instance problems based on hyperelasticity, for which the Expression Rewriter produces a significantly-large number of invariant scalar terms, which dramatically impact the execution time.

We have implemented a procedure that vector-expands scalar expressions, while lifting them outside of the loop nest as shown in Figure~\ref{fig:vectexpanded-code}. The dependency graph is queried to assess the legality of the transformation. At the price of extra memory, vector-expanded terms can now be vectorized by the compiler.

% Expose vectorizability of the outermost loop 
% Render the loop nest perfect

@@

The algorithm is shown in Figure~\ref{fig:algo-dgemm}. By using the rewrite rules in Figure~\ref{fig:rewrite-rules}, it can be proved that the Expression Rewriter can always reduce an assembly expression to a summation, over each quadrature point, of outer products along the test and trial functions. Each outer product is then isolated, i.e. the assembly expression is split into chunks, each chunk representing an outer product. Terms in the bodies of the various enclosing loops (e.g. coefficients evaluation at quadrature point, temporaries introduced by the Expression Rewriter) are vector-expanded and hoisted outside of the loop nest, as explained in Section~\ref{sec:vect-exp}. This implies that the loop nest is now perfect; that is, there is no intervening code among the various loops. The element matrix evaluation became a sequence of perfect loop nests, each loop nest computing a dense matrix-matrix multiply between temporaries hoisted by the Expression Rewriter or tabulated basis functions. Eventually, the storage layout of the involved operands is changed so as to be conforming to the BLAS interface (e.g. two dimensional arrays are flatten as one dimensional arrays). The translation of each loop nest into a DGEMM call is the last, straightforward step.

% 
\subsection{Model-driven Dynamic Autotuning}
We have demonstrated in~\cite{Luporini} that an optimal set of code transformations does not exist. What sequence of transformations applying to a problem to optimize its performance depends on a broad range of factors, including mathematical structure of the input form, polynomial order of employed function spaces, presence of pre-multiplying functions, and, of course, the characteristics of the underlying architecture.

>>>>

\begin{figure}[t]
\tiny
\centering     
\subfigure[Original code]{\label{fig:original-code}\lstinputlisting{listings/simple.code}}
~~~~~~~~~~~~~~~~
\subfigure[Invariant code]{\label{fig:invariant-code}\lstinputlisting{listings/invariant.code}}
~\\
~\\
\subfigure[Factorized code]{\label{fig:factorized-code}\lstinputlisting{listings/factorized.code}}
~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\subfigure[Scalar-expanded code]{\label{fig:scalarexp-code}\lstinputlisting{listings/vectexpanded.code}}
\caption{Original, invariant, factorized, and then scalar-expanded code.}
\end{figure}

Consider the element matrix computation in Figure~\ref{fig:original-code}, which is an excerpt from a Burgers problem. The assembly expression, produced by the FEniCS Form Compiler, has been deliberately simplified, and code details have been omitted for brevity and readability. In practice, assembly expressions can be much more complex, for example depending on the differential operators employed in the variational form; however, this example is representative enough for highlighting patterns that are common in a large class of problems.

A first glimpse of the code suggests that the sub-expression \texttt{a*f0*A[i][j]+b*f1*B[i][j]} is invariant with respect to the innermost (trial functions) loop \texttt{k}, so it can be hoisted at the level of the outer loop \texttt{j} to avoid redundant computation. This is indeed a standard compiler transformation, supported by any available compilers, so, in principle, there should be no need to transform the source code explicitly. With a closer look we notice that the sub-expression \texttt{d*D[i][k]+e*E[i][k]} is also invariant, although, this time, with respect to the outer (test functions) loop \texttt{j}. In~\cite{Luporini}, we have showed that available compilers limit the search for code motion opportunities to the innermost loop of a given loop nest. Moreover, the lack of cost models to ascertain both the optimal place where to hoist an expression and whether or not vectorizing it at the price of extra temporary memory is a fundamental limiting factor. The \textit{generalized loop-invariant code motion} technique that we implemented in COFFEE, leading to the code in Figure~\ref{fig:invariant-code}, addressed these limitations, which are critical in the case of non-trivial assembly expressions, resulting in run-time improvements up to 3$\times$.

In this paper, we target particularly complex variational forms for which a more aggressive transformation is required. We generalize the problem of finding and hoisting invariant sub-expressions by formulating the following critical question: \textit{how an assembly expression should be rewritten so as to minimize both floating point operations and data movement}? To minimize floating point operations, in terms of instructions executed, it is important to reduce the absolute number of arithmetic operations to evaluate the element matrix as well as to achieve full vectorization of the assembly loop nest.

We start considering the case of assembly expressions that, after generalized loop-invariant code motion has been applied, still  ``hide'' opportunities for code hoisting. By examining again the code in Figure~\ref{fig:invariant-code}, we notice that the basis function array \texttt{A} iterating along the \texttt{[i,j]} loops appears twice in the expression. By expanding the products in which \texttt{A} is accessed and applying sum commutativity, we can factorize the expression. This has two effects: firstly, it reduces the number of arithmetic operations performed; secondly, and most importantly, it exposes a new sub-expression \texttt{A[i][k]/c+T2[k]*f} invariant with respect to loop \texttt{j}. Consequently, hoisting can be performed, resulting in the code in Figure~\ref{fig:factorized-code}. In general, exposing factorization opportunities requires traversing the whole expression tree, and then expanding and moving terms. It also needs heuristics to select a factorization strategy: there may be different opportunities of reorganizing sub-expressions, and, in our case, the best is the one that maximizes the invariant code eventually disclosed. We will discuss this aspect formally in Section~\ref{fig:rewrite-rules}.

As a second observation, we note that integration-dependent expressions are inherently executed as scalar code. For example, the \texttt{f0*a} and \texttt{f1*b} products in Figure~\ref{fig:factorized-code} depend on the loop along quadrature points; these operations are performed in a non-vectorized way at every \texttt{i} iteration. This is not an issue in our running example, where scalar computation represents a small percentage of the total, but it becomes a concrete problem in complicated forms, like those at the heart of hyperelasticity (which we evaluate in Section~\ref{sec:perf-results}). In such forms, the amount of computation independent of both test and trial functions loops is so large that it has a significant impact on the run-time, despite being executed only $O(I)$ times (with $I$ number of quadrature points). We have therefore implemented an algorithm to scalar-expand integration-dependent expressions, which leads to codes as in Figure~\ref{fig:scalarexp-code}.

The third point concerns the register pressure induced by the assembly expression. Once the code has been optimized for arithmetic intensity, it is important to think about how the transformations impacted register allocation. Assume the local assembly kernel is executed on a state-of-the-art CPU architecture having 16 logical registers, e.g. an Intel Haswell. Each value appearing in the expression is loaded and kept in a register as long as possible. In Figure~\ref{fig:scalarexp-code}, for instance, the scalar value \texttt{g} is loaded once, whereas the term \texttt{det*W[i]} is precomputed and loaded in a register at every $i$ iteration. This implies that at every iteration of the \texttt{jk} loop nest, 12$\%$ of the available registers are spent just to store values independent of test and trial functions loops. In more complicated expressions, the percentage of registers destined to store such constant terms can be even higher. Registers are, however, a precious resource, especially when evaluating compute-intensive expressions. The smaller is the number of available free registers, the worse is the instruction-level parallelism achieved: for example, a shortage of registers can increase the pressure on the L1 cache (i.e. it can worsen data locality), or it may prevent the effective application of standard transformations, e.g. loop unrolling. We aim at relieving this problem by suitably expanding terms and introducing, where necessary, additional temporary values. We illustrate this in the following example.

@@

\begin{figure}[t]
\tiny
\centering
\subfigure[Expandable code]{\label{fig:toexpand-code}\lstinputlisting{listings/toexpand.code}}
~\\
\subfigure[Expanded 1 code]{\label{fig:expanded-1-code}\lstinputlisting{listings/expanded-1.code}}
~~
\subfigure[Expanded 2 code]{\label{fig:expanded-2-code}\lstinputlisting{listings/expanded-2.code}}
\caption{Expanding code.}\label{fig:expanded-code}
\end{figure}

Consider a variant of the Burgers local assembly kernel, shown in Figure~\ref{fig:toexpand-code}. This is still a representative, simplified example. We can easily distribute \texttt{det*W[i]} over the three operands on the left-hand side of the multiplication, and then absorb it in the pre-computation of the invariant sub-expression stored in \texttt{T1}, resulting in code as in Figure~\ref{fig:expanded-1-code}. Freeing the register destined to the constant \texttt{g} is less straightforward: we cannot absorb it in \texttt{T1} as we did with \texttt{det*W[i]} since the same array is accessed in \texttt{(T1[j]*A[i][k])}. The solution is to add another temporary as in Figure~\ref{fig:expanded-2-code}. Generalizing, this is a problem of data dependencies: to solve it, we use a dependency graph in which we add a direct edge from identifier \texttt{A} to identifier \texttt{B} to denote that the evaluation of \texttt{B} depends on \texttt{A}. The dependency graph is initially empty, and it is updated every time a new temporary is created by either loop-invariant code motion or expansion of terms. The dependency graph is then queried to understand when expansion can be performed without resorting to new temporary values. This aspect is formalized in the next section.

\subsection{Rewrite Rules}
\label{sec:rewrite-rules}
In general, assembly expressions produced by automated code generation can be much more complex than those we have used as examples, with dozens of terms involved (basis function arrays, derivatives, coefficients, ...) and hundreds of (nested) arithmetic operations. Our goal is to establish a portable, unified, platform-independent, and systematic way of reducing the computational strength of an expression exploiting the intuitions described in the previous section. This \textit{expression rewriting} should be simple; definitely, it must be robust to be integrated in an optimizing domain-specific compiler capable of supporting real problems. In other words, we look for an algorithm capable of transforming a plain assembly expression by applying 1) generalized loop-invariant code motion, 2) non-trivial factorization and re-association of sub-expressions, and 3) expansion of terms; after that, it should perform scalar-expansion to achieve full vectorization of the assembly code.

We centre such an algorithm around a set of rewrite rules. These rules drive the transformation of an expression, prescribe where invariant sub-expressions will be moved (i.e. at what level in the loop nest), and track the propagation of data dependencies. When applying a rule, the state of the loop nest must be updated to reflect, for example, the use of a new temporary and new data dependencies. We define the state of a loop nest as $L = (\sigma, G)$, where $G = (V, E)$ represents the dependency graph, while $\sigma : Inv \rightarrow S$ maps invariant sub-expressions to identifiers of temporary arrays. The notation $\sigma_i$ refers to invariants hoisted at the level of loop $i$. We also introduce the \textit{conditional hoister} operator $[]$ on $\sigma$ such that
\begin{gather*}
\sigma[\sfrac{v}{x}] = 
\begin{cases}
\sigma(x) \text{~~~~~~~~if $x \in Inv$; $v$ is ignored}\\
v \text{~~~~~~~~~~~~~if $x$ $\notin$ $Inv$; $\sigma$(x) = v}\\
\end{cases}
\end{gather*}
That is, if the invariant expression \texttt{x} has already been hoisted, $[]$ returns the temporary identifier hosting its value; otherwise, \texttt{x} is hoisted and a new temporary $v$ is created. There is a special case when $v = \perp$, used for \textit{conditional deletion} of entries in $\sigma$. Specifically
\begin{gather*}
\sigma[\sfrac{\perp}{x}] = 
\begin{cases}
\sigma(x) \text{~~~~~~~~if $x \in Inv$; $\sigma = \sigma \setminus (x, \sigma(x))$}\\
v \text{~~~~~~~~~~~~~~if $x$ $\notin$ $Inv$; $v \notin S$}\\
\end{cases}
\end{gather*}
In other words, the invariant sub-expression \texttt{x} is removed and the temporary identifier that was hosting its value is returned if \texttt{x} had been previously hoisted; otherwise, a fresh identifier $v$ is returned. This is useful to express updates of hoisted invariant sub-expressions when expanding terms.

In the following, a generic (sub-)expression is represented with roman letter \texttt{a, b, ...}; constant terms are considered a special case, so greek letters \texttt{$\alpha$, $\beta$} are used instead. The iteration vector $i = [i_0, i_1, ...]$ is the ordered sequence of the indices of the loops enclosing an (sub-)expression. We will refer to $i_0$ as the outermost enclosing loop. The notation $a_i$, therefore, indicates that the expression $a$ assume distinct values while iterating along the loops in $i$; its outermost loop is silently assumed to be $i_0$.

Rewrite rules for expression rewriting are provided in Figure~\ref{fig:rewrite-rules}; obvious rules are omitted for brevity. The Expression Rewriter applies the rules while performing a depth-first traversal of the assembly expression tree. Given an arithmetic operation between two sub-expressions (i.e. a node in the expression tree), we first need to find an applicable rule. There cannot be ambiguities: only one rule can be matched. If the preconditions of the rule are satisfied, the corresponding transformation is performed; otherwise, no rewriting is performed, and the traversal proceeds. As examples, it is possible to instantiate the rules in the codes shown in Figures~\ref{fig:original-code} and~\ref{fig:toexpand-code}; eventually, the optimized codes in Figures~\ref{fig:factorized-code} and~\ref{fig:expanded-2-code} are obtained, respectively.

@@

To what extent should rewrite rules be applied is a question that cannot be answered in general. In some problems, a full rewrite of the expression may be the best option; in other cases, on the other hand, an aggressive expansion of terms, for example, may lead to high register pressure in the loops computing invariant terms, worsening the performance. In Section~\ref{sec:autotune} we explain how to leverage the Code Specializer to select a suitable rewriting strategy for the problem at hand.

@@

@@

The algorithm is shown in Figure~\ref{fig:algo-dgemm}. By using the rewrite rules in Figure~\ref{fig:rewrite-rules}, it can be proved that the Expression Rewriter can always reduce an assembly expression to a summation, over each quadrature point, of outer products along the test and trial functions. Each outer product is then isolated, i.e. the assembly expression is split into chunks, each chunk representing an outer product. Terms in the bodies of the various enclosing loops (e.g. coefficients evaluation at quadrature point, temporaries introduced by rewriting the assembly expression) are vector-expanded and hoisted outside of the loop nest, as explained in Section~\ref{sec:vect-exp}. This implies that the loop nest is now perfect; that is, there is no intervening code among the various loops. The element matrix evaluation has now become a sequence of perfect loop nests, each loop nest computing a dense matrix-matrix multiply between temporaries hoisted by the Expression Rewriter or tabulated basis functions. Eventually, the storage layout of the involved operands is changed so as to be conforming to the BLAS interface (e.g. two dimensional arrays are flatten as one dimensional arrays). The translation of each loop nest into a DGEMM call is the last, straightforward step.

% 
\subsection{Model-driven Dynamic Autotuning}
\label{sec:autotune}
We have demonstrated in~\cite{Luporini} that an optimal set of code transformations does not exist. What sequence of transformations applying to a problem to optimize its performance depends on a broad range of factors, including mathematical structure of the input form, polynomial order of employed function spaces, presence of pre-multiplying functions, and, of course, the characteristics of the underlying architecture.
